---
title: "Predict Movie rating with MovieLens data"
author: Partha Sarathi Mukherjee
date: 5/12/2020
output: pdf_document
---
# Executive Summary
Predicting a movie rating has huge financial impact. A good predictive rating will mean that the movie should be
released in a massive scale with a big upfront advertisement budget. So it is essential that we get rating with
 high accuracy.

The MovieLens data set is a comprehensive dataset thou for this illustration we will limit the dataset to 10 million.

The dataset comprises of the following attributes

 - Movie details like (MovieId, Title)

 - Movie attributes like (Generes, timestamp)

 - Movie viewers(userId)

 - Movie ratings(rating)


A movie can belong in 1 or more of 19 different Genre.

Using the dataset provided, the objective is to build a machine learning model that can predict future movie ratings.

Key steps performed to build movie rating predictor machine learning model are:-

 1. Data Visualization -- How the rating data is distributed among the attributes?

 2. Data Cleanup -- Bad and missing data are trimmed.

 3. Build Models -- Use Matrix Factorization to build recomendation models.

 4. Model Predictions -- Models are evaluated against Test sets

 5. Evaluate Models -- Target is to build a model with RMSE < 0.86490


\pagebreak

# Analysis
There 2 methods succeeded in the original Netflix challenge

 1. Matrix Factorization

 2. Ensemble Model

In this task we use Matrix factorization. It is better suited for recommendation system and gives better RMSE with fewer
 variables.

## Overall data distribution
```{r loading-libs, message=FALSE}
library(tidyverse)
library(dplyr)
```

```{r data-stats, echo=F}
load("ml-10M100K/edx.Rda")
print("Dimensions")
dim(edx)
print("Summary")
str(edx)
print("Summary of distinct users and movies")
summarize(edx, users = n_distinct(userId), movies = n_distinct(movieId))
```

\pagebreak

## Distribution of rating per movie
```{r rating-per-movie, echo=F}
edx %>%
     dplyr::count(movieId) %>%
     ggplot(aes(n)) +
     geom_histogram(bins = 30, color = "black") +
     scale_x_log10() +
     ggtitle("Movies")
```

### Conclusion: Remove movies not rated by atleast 25 users.

\pagebreak

## Distribution of rating per user
```{r rating-per-user, echo=F}
edx %>%
     dplyr::count(userId) %>%
     ggplot(aes(n)) +
     geom_histogram(bins = 30, color = "black") +
     scale_x_log10() +
     ggtitle("Users")

```

### Conclusion: Remove users not rating more than 25 movies.


## Data Abnormality
On denormalization of the data and looking into all the genre, we found that only one movie had no rating (Pull My Daisy
 (1958)). The movie id is 8606.

### Conclusion: Remove movie "Pull My Daisy" (1958) from the dataset.

\pagebreak

# Data Cleaning
Based on above conclusions, data is cleaned as

1. Remove users rating less than 25 movies -- Removed 7551 users. 155124 movies removed. 2.29% decrease

2. Remove movies not rated by atleast 25 users + movie id 8606 -- Removed 30576 movies. 0.4 % decrease.

Cleaned Data now looks like:
```{r cleaned-data-stats, echo=F}
load("ml-10M100K/edx_clean.Rda")
print("Dimensions")
dim(edx_clean)
print("Summary")
str(edx_clean)
```

\pagebreak

# Modeling -- Martix Factorization
Steps to build Matrix Factorization model as:-

## Step 1: Average Movie Rating
Build an model which predicts the same rating (average movie rating) regardless of the user.

## Step 2: Add Movie Effect on Rating
Compute the movie effect based on the average difference between the actual rating and the average rating.
Build a model with average movie rating from step 1 and movie effect value.

## Step 3: Add User Effect on Rating
Compute the user effect based on the average difference between actual rating and the sum of average and movie rating obtained in step 2.
Build a model with average movie rating from step 1, movie effect value from step 2 and user effect from this step.

# Result
Using the above 3 steps, the final model RMSE dropped to 0.862.

Best performance requirement for this problem is achieved (RMSE < 0.86490).

```{r rmse-results, echo=F}
load("ml-10M100K/rmse_results.Rda")
knitr::kable(rmse_results)
```

# Conclusion
The Matrix Factorization method is the best approach to recommendation systems.

More steps can added to improve the solution like

 1. User's Age: Amount of time user is in system from user's first rating,
 2. Movie's Age: Amount of time passed from the first rating.
 3. Movie's Expectation: How many people already rated the movie

 and so on...

